hatchR: A toolset to predict hatch and emergence phenology in wild fishes

### 2025-06-25 inst/manuscript/paper.Rmd
Add i.e. to parenthetical
Remove Table 1

As poikilothermic organisms, fish have developmental rates that are strongly influenced by ambient environmental conditions. This close relationship has allowed researchers to develop statistical models that predict developmental phenology with high accuracy. Historically, these models were created in aquaculture settings under the assumption of constant temperatures throughout development [@beacham1990; @mcphail1979; @alderdice1978]. As a result, accumulated thermal unit thresholds (ATU models) have become widely used to predict development in wild populations. However, this approach can be inaccurate, particularly at the cold and warm extremes of temperature ranges, due to the non-linear relationship between average incubation temperature and the thermal units required for development [@quinn2018].

@@

The **hatchR** software assumes input data consists of at least two required columns: a date column indicating the date (and optionally time, as is commonly the case for field loggers) of each temperature measurement, and a temperature column providing the corresponding temperature measurement (in °C). Other columns may be present, but column names should not include spaces. Data should follow the format outlined in Table 1.

@@

# Table Captions

**Table 1.** Example temperature data for use in hatchR.

>>>>

As poikilothermic organisms, fish have developmental rates that are strongly influenced by ambient environmental conditions. This close relationship has allowed researchers to develop statistical models that predict developmental phenology with high accuracy. Historically, these models were created in aquaculture settings under the assumption of constant temperatures throughout development [@beacham1990; @mcphail1979; @alderdice1978]. As a result, accumulated thermal unit thresholds (*i.e.*, ATU models) have become widely used to predict development in wild populations. However, this approach can be inaccurate, particularly at the cold and warm extremes of temperature ranges, due to the non-linear relationship between average incubation temperature and the thermal units required for development [@quinn2018].

@@

The **hatchR** software assumes input data consists of at least two required columns: a date column indicating the date (and optionally time, as is commonly the case for field loggers) of each temperature measurement, and a temperature column providing the corresponding temperature measurement (in °C). Other columns may be present, but column names should not include spaces.

@@
